Millions of Type Errors

### 2023-07-31 programming23/paper.tex
telemetry design notes

\overfullrule=1mm

@@

Beyond their immediate applications to {Lua},
answers to these questions have implications for gradual types,
success types, and semantic subtyping.

@@

  %% FILL examples for each error?
  \begin{tabular}{ll}
    Label & Interpretation \\\midrule
    \code{CodeTooComplex} & Type analysis failed, cannot understand the code \\
    \code{UnificationTooComplex} & Type analysis failed, unification solver hit a limit \\
    \code{SyntaxError} & Basic parse error, e.g., \code{for if end} \\

    \code{CountMismatch} & Arity mismatch for a function \\
    \code{IncorrectGenericParameterCount} & Arity mismatch for a generic type \\
    \code{UnknownProperty} & Referenced an invalid field or method  \\
    \code{OnlyTablesCanHaveMethods} & Tried to attach a method to a non-table \\
    \code{CannotCallNonFunction} & Called a value that is not a function \\
    \code{TypesAreUnrelated} & Failed to cast, unify, or check subtyping \\
    \code{TypeMismatch} & Generic label for other type errors \\
    \code{GenericError}
    & Generic label for other non-type errors, such as looping\!\!\! \\
    & over a table without a clear iteration strategy
    %% & {}~~declaring a supertype that is not a class type

@@

No PII whatsoever

Random session selection

Random events during session, except module switch.

No access to other key events: save, exit.
Don't want access to run, publish.

>>>>

%%\overfullrule=1mm

@@

Beyond their immediate applications to {Lua},
answers to these questions have implications for the design
of gradual types, success types, and semantic subtyping.
Luau represents the first large-scale combination of these features,
and lessons from this experience can inform future applications.

@@

  %% FILL examples for each error?
  \begin{tabular}{ll}
    Label & Interpretation \\\midrule
    \code{CodeTooComplex} & Type analysis failed, cannot understand the code\!\!\! \\
    \code{UnificationTooComplex} & Type analysis failed, unification solver hit a limit\!\!\! \\
    \code{SyntaxError} & Basic parse error, e.g., \code{for if end} \\

    \code{CountMismatch} & Arity mismatch for a function \\
    \code{IncorrectGenericParameterCount} & Arity mismatch for a generic type \\
    \code{UnknownProperty} & Referenced an invalid field or method  \\
    \code{OnlyTablesCanHaveMethods} & Tried to attach a method to a non-table \\
    \code{CannotCallNonFunction} & Called a value that is not a function \\
    \code{TypesAreUnrelated} & Failed to cast, unify, or check subtyping \\
    \code{TypeMismatch} & Generic label for other type errors \\
    \code{GenericError}
    & Generic label for other non-type errors, such as\\
    & \hbox{}~~looping over an unordered table
    %% attempting to extend a type that does not describe a class
    %% attempting to iterate over a table without a clear order

@@

\paragraph{Goals:}

\begin{itemize}
  \item
    Learn about type system adoption and use
  \item
    Discover roadblocks to adoption
  \item
    Assess the helpfulness of error highlights
\end{itemize}

\paragraph{Limits:}

\begin{itemize}
\item
  No personal data (PII)
  \subitem
    Fairly obvious.
    But, we also can't collect directory names, filenames, or code snapshots
    because these all might have PII throughout.
\item
  No data about public events.
  \subitem
    Contributions to the Roblox Marketplace, for instance, are a public event.
    Although it would be interesting to know the number of type errors
    in published experiences, this data might be used to identify creators
    who publish often --- or on a distinct schedule.
\item
  Low cost
  \subitem
    Building, sending, and storing the data must all run quickly.
    FILL alan: what was our hard cap on Kibana use?
\end{itemize}

\paragraph{Solutions:}

\begin{itemize}
  \item
    Collect metrics instead of source code: lines, modules, num errors
  \item
    Track an edit range to compare type errors with edits,
    use a coarse $[min, max]$ interval for speed
  \item
    Count specific errors only for the current edit range
    (reduces size of logs)
  \item
    Log randomly on keystroke instead of per specific event.
    One exception: log module switches because that helps discover
    codebases that use several analysis modes.
    Randomness helps lower the amount of data.
  \item
    Observe sessions randomly (again to lower the amount of data).
\end{itemize}

%% Extra Roblox constraints:
%% - No access to some events: save, exit
%% - No idea if analysis widget is open, closed, or obscured
%% - Don't know which module is the data model
%%   (no ids for ANY module for that matter)
